---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eq5d.plus

<!-- badges: start -->
<!-- badges: end -->

The goal of eq5d.plus is to run linear models on EQ-5D quality of life data that uses both the score and visual analogue scale (VAS).

## Installation

You can install the released version of eq5d.plus from [GitHub](https://github.com/) with:

```r
# use the devtools library
library(devtools)
# Install the latest version from GitHub:
install_github("agbarnett/lmvas")
```


## Example

This example runs a linear model with independent variables of age and sex:

```{r example}
library(eq5d.plus)
## example using the Spanish osteoarthritis data
model = lmvas(vas_name='EQ5D_VAS', eq_name='EQ5D', independent_vars = c('age','sex'), data=arthritis)
summary(model)
```

The results show a reduced intercept for the VAS compared with the score of `r round(model$summary.fixed$mean[2]*1000)/1000`. 
Quality of life is lower on average for women and for older respondents.

The score and VAS are positively correlated with an estimated correlation of `r round(model$summary.hyperpar$mean[4]*1000)/1000` ("rho" estimate in model hyperparameters).

The model may take a short while to run for very large data sets.
